---
output:
  github_document:
    html_preview: false
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r knitrsetup, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
options(tibble.print_min = 5, tibble.print_max = 5)
```

```{r mainexample, echo = FALSE}
suppressMessages(suppressWarnings(library(bbqr)))
```

[![CRAN](http://www.r-pkg.org/badges/version/bbqr)](http://cran.r-project.org/package=bbqr) [![Travis build status](https://travis-ci.org/wlandau/bbqr.svg?branch=master)](https://travis-ci.org/wlandau/bbqr) [![AppVeyor build status](https://ci.appveyor.com/api/projects/status/github/wlandau/bbqr?branch=master&svg=true) [![Codecov](https://codecov.io/github/wlandau/bbqr/coverage.svg?branch=master)](https://codecov.io/github/wlandau/bbqr?branch=master)

# A bare-bones message queue for R

The `bbqr` package is a way for parallel processes in R to send messages to each other. Let's say Process A and Process B are working on a parallel task together. First, both processes grab the queue.

```{r setup}
path <- tempfile() # Define a path to your queue.
path # In real life, temp files go away when the session exits, so be careful.
q <- bbqr(path) # Create the queue.
```

The queue uses flat text files in the folder you specified.

```{r files}
list.files(q$path) # The queue lives in this folder.
q$list() # You have not pushed any messages yet.
q$empty()
q$count()
```

Then, Process A sends instructions to Process B.

```{r ab}
q$push(title = "Hello", message = "process B.")
q$push(title = "Calculate", message = "sqrt(4)")
q$push(title = "Calculate", message = "sqrt(16)")
q$push(title = "Send back", message = "the sum.")
```

You cna inspect the contents of the queue from either process.

```{r list}
q$list()
q$count()
q$empty()
```

As Process A is pushing the messages, Process B can consume them.

```{r pop2}
q$pop(2) # If you pass 2, you are assuming the queue has >=2 messages.
```

Those "popped" messages are not technically in the queue any longer, but we can still see a full log of all the messages that were ever sent.


```{r listafterpop}
q$list()
q$list(1) # You can specify the number of messages to list.
q$log()
```

Let's let Process B get the rest of the instructions.


```{r bfinishconsume}
q$pop() # q$pop() with no arguments just pops one message.
q$pop()
```

Now let's say Process B follows the instructions in the messages. The last step is to send the results back to Process A.

```{r sendback}
q$push(title = "Results", message = as.character(sqrt(4) + sqrt(16)))
```

Process A can now see the results.

```{r aconsume}
q$pop()
```

When you are done, you have the option to destroy the files in the queue.

```{r destroy}
q$destroy()
file.exists(q$path)
```

This entire time, the queue was locked when either process was trying to create, access, or modify it. That way, the results stay correct even when multiple processes try to read or change the data at the same time.

# Similar work

## liteq

[Gábor Csárdi](https://github.com/gaborcsardi)'s [`liteq`](https://github.com/r-lib/liteq) package offers essentially the same functionality implemented with SQLite databases. It has a few additional features (for example, the ability to detect crashed workers and requeue failed messages). However, at the time `bbqr` was implemented, `liteq` was [still in an early stage of development](https://github.com/r-lib/liteq/issues/17).

## Other message queues

There is a [plethora of message queues](http://queues.io/) beyond R, most notably [ZeroMQ](http://zeromq.org) and [RabbitMQ](https://www.rabbitmq.com/). In fact, [Jeroen Ooms](http://github.com/jeroen) and [Whit Armstrong](https://github.com/armstrtw) maintain [`rzmq`](https://github.com/ropensci/rzmq), a package to work with [ZeroMQ](http://zeromq.org) from R. These tools may be ideal for intermediate advanced users, but `bbqr` has two main advantages for simple use cases.

1. It does not require you to install anything outside R.
2. Tools based on IP/TCP sockets may only be able to send one message at a time, but `bbqr` can send multiple messages before any are consumed. Some applications may need to allow for a backlog of unread messages.
